### 2021-05-28 cued/additional.tex
Copy over relevant bits

For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}}(b, e, k)
\end{equation}
%
With edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$. The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Omega(e)$. This is why this formulation is given the microcanonical moniker. If we consider the half-edges to be distinguishable for a moment, the total number of configurations that satisfy the $e$ constraint is:
%
\begin{equation}
	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!}
\end{equation}
%
Where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!! \coloneqq 2^m m!$. Nevertheless, a great number of these configurations yield the same graph $A$. We denote the number of configurations that yield the adjacency matrix $A$ with $\Xi(A)$, which can be computed as:
%
\begin{equation}
	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }
\end{equation}
%
Note the similarity between the forms of $\Omega(e)$ and $\Xi(A)$ as $e$ is effectively the adjacency matrix of the block-graph. With these defined we can write the overall likelihood:
%
\begin{equation}
	p(A|k,e,b) = \frac{\Xi(A)}{\Omega(e)}
\end{equation}
%
Obviously, this form is only defined if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0. With the likelihood defined, we move on to the prior. As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM and not a parameter so we are not free to choose a prior for it. Nevertheless, we can borrow the conditional prior proposed by \citet{Peixoto-Bayesian-Microcanonical} for the other parameters to give:
%
\begin{equation}
	p(e, k | b) = p(e | b) p(k| e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
\end{equation}
%
Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$. $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph. Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$. The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$. Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$. $q(m, n)$ is related to but different from $\specialchoose{n}{m}$. Recall that $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.

The form of these priors were chosen carefully by \citet{Peixoto-Bayesian-Microcanonical} to more closely match the structure of empirical networks than simple uniform priors. We do not repeat his arguments here.

@@

For sampling from the $\theta$-chain of the block membership generator parameters, we employ the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
%
\begin{equation}
	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
\end{equation}
%
There are two competing objectives when choosing the step-size $h_t$. On the one hand, we want the step-size to be large so that we arrive at a high density region quickly. However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling. A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing. Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$. As well as the trivial constraint for $h_t$ to be strictly positive, we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
%
\begin{equation}
	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
	\sum_{t=1}^{\infty} h_t^2 < \infty
	\label{eqn:h-constraints}
\end{equation}
%
The first constraint ensures that we can cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point. The second constraint ensures that once we converge to the mode we will stay there, rather than simply bouncing around it. \citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
%
\begin{equation}
	h_t = \alpha(\beta + t)^{-\gamma}
\end{equation}
%
Where $\alpha, \beta, \gamma$ are hyper-parameters to be chosen. We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}. To reduce the number of hyperparameters we set these to have values given by the equations in \ref{eqn:step-size-params}.
%
\begin{equation}
	\alpha = \frac{250 \cdot s}{N} \qquad \beta = 1000 \qquad \gamma = 0.8
	\label{eqn:step-size-params}
\end{equation}
%
Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling. For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation. If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}. This speeds up computation at the expense of the exactness of the method. Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability is close 1.

Nevertheless, to choose the value of $s$ we must quantify the trade-off between acceptance ratio and burn-in time. We do this by way of an example with the primary school dataset \cite{schools}. We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\Tcal_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$. We then use this to run the $\theta$-chain for $T_\theta = 10,000$ iterations. The acceptance ratio -- which we denote $r_\alpha$ -- is easy enough to compute as simply the fraction of proposed moves which we accept. However, to quantify burn-in time we must use a different metric. We can plot the mean of the objective function averaged over all samples as a proxy for burn-in time:
%
\begin{equation}
	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right)
\end{equation}
%
As the chain will equilibrate in the vicinity of a minima in $U(\theta)$, the average over all samples is a rough indicator of the speed of burn-in. This assumes that the random starting point has high $U(\theta)$. This assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ simply from a lucky guess. The higher the value of $\bar{U}$, the longer we infer the chain took to burn in and reach equilibrium. We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on figure \ref{fig:step-size-tradeoff}.
%
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.6\linewidth]{step-size-tradeoff}
	\caption{Primary school dataset \cite{schools} illustration of step-size tradeoff}
	\label{fig:step-size-tradeoff}
\end{figure}

@@

As with any MCMC method, we must deal with the issues presented by burn-in and thinning. We have introduced the notation $\Tcal_b$ and $\Tcal_\theta$ to denote the set of samples we keep from the $b$ and $\theta$ chains respectively. Note that we generate $T_b$ and $T_\theta$ samples total. The burn-in period refers to the time taken for the Markov Chain to converge to the stationary distribution. Sample thinning is necessary to ensure that neighbouring samples satisfy independence. However, as we do not leverage the independence property this is less important in our analysis. We can write the general set $\Tcal_\star$ as:
%
\begin{equation}
	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}
\end{equation}
%
Where the parameter $\kappa_\star \in (0, 1)$ controls our burn-in and $\lambda_\star$ controls our thinning. $\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ or $U(\theta^{(t)})$ with respect to the epoch $t$. $\kappa_\star$ is then chosen to encompass the region where the log-target has roughly equilibrated. As we do not leverage sample independence $\lambda_\star$ can be chosen less rigorously. We often just use $\lambda_b=5$ and $\lambda_\theta = 10$.

@@

For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist. We could however, allow our choice of $B$ to be influenced by the observed data. This places us in the domain of empirical Bayes, which must be negotiated carefully. Prior beliefs must be determined a priori else they are not prior. However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary. Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.

If we allow $B$ to vary in the $b$-chain (i.e. new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached. We take the number of non-empty blocks at the MDL solution to be our fixed block number $B$ for subsequent analysis. Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.

>>>>

For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}}(b, e, k)
\end{equation}
%
With edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$. The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Omega(e)$. This is why this formulation is given the microcanonical moniker. If we consider the half-edges to be distinguishable for a moment, the total number of configurations that satisfy the $e$ constraint is:
%
\begin{equation}
	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!}
\end{equation}
%
Where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!! \coloneqq 2^m m!$. Nevertheless, a great number of these configurations yield the same graph $A$. We denote the number of configurations that yield the adjacency matrix $A$ with $\Xi(A)$, which can be computed as:
%
\begin{equation}
	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }
\end{equation}
%
Note the similarity between the forms of $\Omega(e)$ and $\Xi(A)$ as $e$ is effectively the adjacency matrix of the block-graph. With these defined we can write the overall likelihood:
%
\begin{equation}
	p(A|k,e,b) = \frac{\Xi(A)}{\Omega(e)}
\end{equation}
%
Obviously, this form is only defined if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0. With the likelihood defined, we move on to the prior. As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM and not a parameter so we are not free to choose a prior for it.

Here we recall for reference
the priors $p(\psi | b)$ from \cite{Peixoto-Bayesian-Microcanonical}:
%
\begin{equation}
	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(\psi_k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right],
\end{equation}
%
where $\specialchoose{n}{m}$ is shorthand 
for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$,
which can be thought of as the total number of distinct histograms 
produced by $m$ samples in $n$ bins.
The value
$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph. 
Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform in $e$.
The variable $\eta_j^r$ denotes the number of vertices in block $r$ 
that have degree $j$; formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. 
The denominator $q(m, n)$ denotes the number of different histograms 
produced by $m$ samples in 
at most $n$ non-zero bins that sum to $m$. 
Finally, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number 
of half edges in block $r$ and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ 
is the number of vertices assigned to block $r$.

These were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
more closely match the structure of empirical networks than simple 
uniform priors. We do not repeat these arguments here.

@@

Recall that in 
Section~\ref{s:sfb} we used 
the Metropolis Adjusted Langevin Algorithm (MALA)
to 
sample from the $\theta$-chain of the block membership 
generator parameters.
At iteration $t$, the proposed sample is generated by:
%
\begin{equation}
	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi.
\end{equation}
%
There are two competing objectives when choosing the step-size $h_t$. 
On the one hand, $h_t$ needs to be large so that the sampler
arrives at a high density region quickly,
while too large a step-size would lead to low acceptance rates and thus 
inefficient sampling. An effective strategy is
to use {\em simulated annealing}: allow $h_t$ to slowly decrease
with $t$, as long as $h_t>0$ for all $t$ and also:
%
\begin{equation}
	\sum_{t=1}^{\infty} h_t = \infty, \qquad \textrm{and} \qquad
	\sum_{t=1}^{\infty} h_t^2 < \infty.
	\label{eqn:h-constraints}
\end{equation}
%
Following \citet{Bayesian-SGLD}, we adopt the 
polynomially decaying step-sizes,
%
$h_t = \alpha(\beta + t)^{-\gamma}$,
%
where $\alpha>0$, $\beta>0$ and $\gamma\in(1/2,1]$ are hyper-parameters.
We make the specific choices,
%
\begin{equation}
	\alpha = \frac{250 \cdot s}{N}, \qquad \beta = 1000, \qquad \gamma = 0.8,
	\label{eqn:step-size-params}
\end{equation}
%
where $N$ is the number of data-points and $s$,
the {\em step-size scaling}, is the only free parameter.
%
%For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation. If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}. This speeds up computation at the expense of the exactness of the method. Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability is close 1.
%
Nevertheless, to choose the value of $s$ we must quantify the trade-off between acceptance ratio and burn-in time. We do this by way of an example with the primary school dataset \cite{schools}. We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\Tcal_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$. We then use this to run the $\theta$-chain for $T_\theta = 10,000$ iterations. The acceptance ratio -- which we denote $r_\alpha$ -- is easy enough to compute as simply the fraction of proposed moves which we accept. However, to quantify burn-in time we must use a different metric. We can plot the mean of the objective function averaged over all samples as a proxy for burn-in time:
%
\begin{equation}
	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right)
\end{equation}
%
As the chain will equilibrate in the vicinity of a minima in $U(\theta)$, the average over all samples is a rough indicator of the speed of burn-in. This assumes that the random starting point has high $U(\theta)$. This assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ simply from a lucky guess. The higher the value of $\bar{U}$, the longer we infer the chain took to burn in and reach equilibrium. We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on figure \ref{fig:step-size-tradeoff}.
%
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.6\linewidth]{step-size-tradeoff}
	\caption{Primary school dataset \cite{schools} illustration of step-size tradeoff}
	\label{fig:step-size-tradeoff}
\end{figure}

@@

%As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
When sampling from the $b$ and $\theta$-chains described
in Section~\ref{sec:inference}, we respectively generate
$T_b$ and $T_\theta$ samples total.
We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
as corresponding to a ``burn-in'' period required for the distribution 
of the chain to reach a distribution close to our target,
and we also ``thin'' the remaining samples to 
obtain a less-dependent version. For $\star\in\{b,\theta\}$,
the remaining sample sets are denoted $\Tcal_\star$
in the notation of Section~\ref{s:ss},
%
\begin{equation}
	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
\end{equation}
%
where $\lambda_\star$ controls the thinning. The choice of
$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ 
or $U(\theta^{(t)})$ as a function of $t$,
and choosing $\kappa_\star$ to encompass the region where the log-target has 
roughly reached equilibrium. As we do not leverage sample independence,
$\lambda_\star$ can be chosen less rigorously; we often simply
use $\lambda_b=5$ and $\lambda_\theta = 10$.

@@

For the purposes of the FFBM model, the number of blocks $B$ is a constant 
which must be specified. If the choice of $B$ is influenced 
by the observed data, then the analysis is no longer ``fully Bayesian''
and belongs to the class of methods referred to as ``empirical Bayes.''
However, as the number of blocks only specifies the coarseness of the 
analysis, it is reasonable to allow it to vary. Indeed, 
\citet{peixoto-determine-B} shows that for a fixed 
average degree the maximum number of detectable blocks scales 
as $O(\sqrt{N})$ where $N$ is the number of vertices.

If $B$ is allowed to vary in the $b$-chain (i.e.,
when new blocks can be created and empty blocks are allowed),
then the chain can be run until a minimum description 
length (MDL) solution is reached. We take the number of non-empty blocks 
in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
the necessary burn-in time can be greatly reduced.

### 2021-05-28 cued/derivations.tex
Copy over relevant bits

\subsection{Derivation of conditional block distribution given feature matrix}
\label{appdx:b|x}

We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
%
\begin{align*}
	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
\end{align*}
%
We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$. In other words, the integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$. As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables. Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
%
\begin{align*}
	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
\end{align*}
%
As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution. $\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take. We know $b_i \in [B]$. Therefore, $b \in [B]^N$ and $|[B]^N| = B^N = \nicefrac{1}{c}$. Putting this all together we conclude that:
%
\begin{equation}
	p(b | X) = B^{-N}
\end{equation}
%
This is an extremely important result as it allows samples to be drawn from $p(b | A, X) \propto p(A|b, X) \cdot p(b | X) = p(A|b) \cdot p(b|X)$ without a complicated Monte-Carlo integration over $\theta$ to determine $p(b|X)$. Indeed, we do not even need to know the exact value of $X$ to compute $p(b|X)$. 
\subsection{Derivation of U form}
\label{appdx:form-U}

The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
%
\begin{align}
	\pi_\theta(\theta) &\propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right) \\
	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
\end{align}
%
Where we have introduced $U(\theta)$ equal to the negative log posterior. Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}) by defining $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
%
\begin{equation}
	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
	\label{eqn:U-constituent-terms}
\end{equation}
%
Discarding constant terms, we write $U(\theta)$ as in equation \ref{eqn:U-form-appdx-0}. Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
%
\begin{equation}
	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
	\label{eqn:U-form-appdx-0}
\end{equation}
%

\subsection{Derivation of U gradient with respect to feature parameters}
\label{appdx:gradu}
The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
%
\begin{equation}
	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
	\label{eqn:U-form-appdx}
\end{equation}
%
Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
%
\begin{equation}
	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{r \in [B]} \exp(w_r^T \tilde{x}_i)}
	\label{eqn:a-ij} 
\end{equation}
%
We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$. Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$. To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
%
\begin{align}
	\frac{\partial a_{ij}}{\partial w_k} &= \frac
	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) 
		- 
		\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
\end{align}
%
Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
%
\begin{equation}
	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} ||w_r||^2 \right) = 2w_k
\end{equation}
%
We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
\begin{align}
	\frac{\partial U}{\partial w_k} &= 
	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
\end{align}
%
This is the required result. This form can be computed efficiently through matrix operations. The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.

>>>>

\subsection{Derivation of {\boldmath $p(b|X)$}}
\label{appdx:b|x}

We determine the form of $p(b| X)$ by integrating
out the parameters $\theta$. From the definitions we have:
%
\begin{align*}
	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}.
\end{align*}
%
The key observation here is that 
the value of the integral is independent
of the value of $b_i \in [B]$ as
the integrand has the same form regardless of $b_i$. This is
because the prior is the same for each $w_j$. 
Therefore, the integral is only a function of $\tilde{x}_i$ and $\sigma_\theta^2$,
which means that, as a function of $b$, $p(b|X)\propto 1$. As
$b$ takes values in $[B]^N$, we necessarily have:
%
\begin{equation}
	p(b | X) = \frac{1}{\big|[B]^N\big|}=B^{-N}.
\end{equation}

\subsection{Derivation of {\boldmath $\;U(\theta)$}}
\label{appdx:form-U}

Recall from~(\ref{eq:U}) in Section~\ref{s:sfb} that,
$$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
$$ 
so that $U$ can be expressed as,
$$
U(\theta) 
= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}.
$$
Writing,
$y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
%
\begin{equation}
	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
	\|\theta \|^2,
	\label{eqn:U-constituent-terms}
\end{equation}
%
where
$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j \in [B]} \|w_j\|^2$ 
is the Euclidean norm of the vector of parameters $\theta$.
Therefore, discarding constant terms, we 
obtain exactly the representation~(\ref{eqn:U-form}), as claimed.

\subsection{Derivation of {\boldmath $\;\nabla U(\theta)$}}
\label{appdx:gradu}
Here we show how the gradient 
$\nabla U(\theta)$ can be computed explicitly.
Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}).
Writing $\theta$ as
$\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$,
in order to compute the gradient
$\nabla U(\theta)$ we need to compute
each of its components,
$\nicefrac{\partial U}{\partial w_k}$,
$1\leq k\leq B$.
To that end, we first compute,
%
\begin{align}
	\frac{\partial a_{ij}}{\partial w_k} &= \frac
	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) 
		- 
		\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
	\label{eq:dadw}
\end{align}
%
where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$,
and we also easily find,
%
\begin{equation}
	\frac{ \partial}{\partial w_k} \|\theta\|^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} \|w_r\|^2 \right) = 2w_k.
	\label{eq:dtsdw}
\end{equation}
%
Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
\begin{align}
	\frac{\partial U}{\partial w_k} &= 
	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
	\left( -\frac{\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
	&=  - \left( \sum_{i \in [N]} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
	&= - \left( \sum_{i \in [N]} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
\end{align}
%
This can be computed 
efficiently through matrix operations. The only property of $y_{ij}$ 
we have used in the derivation is the constraint $\sum_{j \in [B]} y_{ij} = 1$,
for all $i$.
